---
title: "Class18"
author: "Sabrina Koldinger (A16368238)"
format: pdf
editor: visual
---

# Class 18

Lets examine and explore Pertusis cases in the US.
URL: https://www.cdc.gov/pertussis/surv-reporting/cases-by-year.html

Use datapasta to scrape this data from the website into R:

```{r}
cdc = data.frame(
                                Year = c(1922L,1923L,1924L,1925L,1926L,
                                         1927L,1928L,1929L,1930L,1931L,
                                         1932L,1933L,1934L,1935L,1936L,1937L,
                                         1938L,1939L,1940L,1941L,1942L,
                                         1943L,1944L,1945L,1946L,1947L,1948L,
                                         1949L,1950L,1951L,1952L,1953L,1954L,
                                         1955L,1956L,1957L,1958L,1959L,
                                         1960L,1961L,1962L,1963L,1964L,1965L,
                                         1966L,1967L,1968L,1969L,1970L,
                                         1971L,1972L,1973L,1974L,1975L,1976L,
                                         1977L,1978L,1979L,1980L,1981L,
                                         1982L,1983L,1984L,1985L,1986L,1987L,
                                         1988L,1989L,1990L,1991L,1992L,1993L,
                                         1994L,1995L,1996L,1997L,1998L,
                                         1999L,2000L,2001L,2002L,2003L,2004L,
                                         2005L,2006L,2007L,2008L,2009L,
                                         2010L,2011L,2012L,2013L,2014L,2015L,
                                         2016L,2017L,2018L,2019L,2020L,
                                         2021L),
        Cases = c(107473,164191,165418,152003,
                                         202210,181411,161799,197371,166914,
                                         172559,215343,179135,265269,180518,
                                         147237,214652,227319,103188,183866,
                                         222202,191383,191890,109873,133792,
                                         109860,156517,74715,69479,120718,
                                         68687,45030,37129,60886,62786,
                                         31732,28295,32148,40005,14809,11468,
                                         17749,17135,13005,6799,7717,9718,
                                         4810,3285,4249,3036,3287,1759,
                                         2402,1738,1010,2177,2063,1623,1730,
                                         1248,1895,2463,2276,3589,4195,
                                         2823,3450,4157,4570,2719,4083,6586,
                                         4617,5137,7796,6564,7405,7298,
                                         7867,7580,9771,11647,25827,25616,
                                         15632,10454,13278,16858,27550,18719,
                                         48277,28639,32971,20762,17972,
                                         18975,15609,18617,6124,2116)
      )
head(cdc)
```

Plot it as cases per year
```{r}
library(ggplot2)
ggplot(cdc) + aes(x=Year,y=Cases) + geom_line() + geom_vline(xintercept =1947, col= "purple") +
  geom_vline(xintercept =1992, col= "red")+
  geom_vline(xintercept =2020, col= "orange")+
  ylab("Cases of Pertussis")
```


Import data from CMI-PB project. This database uses API to return JSON format data. 
We will use `jsonlite`.

```{r}
library(jsonlite)

subject=read_json("http://cmi-pb.org/api/subject", simplifyVector= TRUE)
head(subject)
```


How many people have the wP vs aP (acellular vaccine) in the dataset?
```{r}
table(subject$infancy_vac)
```
60 with aP and 58 with wP

How many male or female are in the study?
```{r}
table(subject$biological_sex)
```
79 women and 39 men

Their races? both combined?
```{r}
table(subject$race)
table(subject$race,subject$biological_sex )
```


ages/ dates

We can use the lubridate package to ease the pain of dealing with dates.
```{r}
library(lubridate)
today()
```

```{r}
today() - ymd("2000-01-01")
today() - ymd("2002-07-01")
```



What is the age of everyone on our dataset?
```{r}
subject$age=round(time_length(today()- ymd(subject$year_of_birth), "years"))
```

```{r}
ggplot(subject)+aes(age, fill=infancy_vac)+geom_histogram()+ facet_wrap(vars(infancy_vac), nrow = 2)
```


More data from CMI-PB

```{r}
specimen= read_json("http://cmi-pb.org/api/specimen", simplifyVector=TRUE)
head(specimen)
```


Use **join** to combine the specimen and subject tables to make a meta table with the metadata. We use the `dplyr` functions for this. 

```{r}
library(dplyr)
meta= inner_join(subject, specimen)
head(meta)
```


Read in other data from CMI-PB
```{r}
ab_titer= specimen= read_json("http://cmi-pb.org/api/v4/plasma_ab_titer", simplifyVector=TRUE)
head(ab_titer)
```

Combining meta and ab_titer with `inner_join()` 
```{r}
alldata= inner_join(meta, ab_titer)
head(alldata)
```

Plot of MFI vs. antigen

```{r}
ggplot(alldata)+ aes(MFI,antigen)+ geom_boxplot()
```


Why are certain antigens and not others very carable in their detected levels here?


Is there a difference between aP and wP? 

```{r}
ggplot(alldata)+ aes(MFI,antigen, col=infancy_vac)+ geom_boxplot()
```
Not a significant difference just by looking. There are some differences but hard to tell. 

```{r}
table(alldata$dataset)
```


Focus in on the 2021 dataset.
```{r}
alldata.21=filter(alldata, dataset== "2021_dataset")
table(alldata.21$dataset)
```

PT antigen IgG levels

```{r}
pt.21= filter(alldata.21, isotype == "IgG",  antigen == "PT")
```



```{r}
ggplot(pt.21) + aes(x=planned_day_relative_to_boost, y=MFI_normalised, col=infancy_vac) +geom_point()
```


```{r}
ggplot(pt.21) + aes(x=planned_day_relative_to_boost, y=MFI_normalised, col=infancy_vac) +geom_line() +geom_point()
```




```{r}
ggplot(pt.21) + aes(x=planned_day_relative_to_boost, y=MFI_normalised, col=infancy_vac, group=subject_id) +geom_line() +geom_point()

```


